---
title: "Homework-3"
author: "Patty Zhang"
date: "11/9/2018"
output: html_document
---

# 1. CASL page 117, question 7.
 Kernels can also be used as density estimators. Specifically, we have
$$f_h(x)=\dfrac{1}{n}\sum_{i}K_h(x-x_i)$$
In this setting we see again why it is important to have the integral of the kernel equal to 1. Write a function kern_density that accepts a training vector x, bandwidth h, and test set x_new, returning the kernel density estimate from the Epanechnikov kernel. Visually test how this performs for some hand constructed datasets and bandwidths.

**answer**:
Start with running the Epanechnikov kernel function with input x, h
Epanechnikov kernel is given by 
$$K(x) = \frac{3}{4} (1-x^2) . 1_{|x| \leq 1}$$
```{r}
kernel_epan <- function(x, h = 1) {
  x <- x / h
  ran <- as.numeric(abs(x) <= 1)
  val <- (3/4) * ( 1 - x^2 ) * ran * (1/h)
  return(val)
}
```

Use the kern_density function to calculate the kernel density with Inputs x, h, and x_new
The function will show how the kernel density estimate by the Epanchnikov kernel

```{r}
kern_density <- function(x, h, x_new) {
  sapply(x_new, function(xi) {
    kh <- kernel_epan((x - xi), h = h)
    den <- sum(kh)/length(x)
    den
  })
}
```

Visually test how this performs for some hand constructed datasets normal distribution N(0,1) and bandwidths with h = 0.01, 0.05, 0.5 and 1.

From the following plots, I learnt that the estimate is smoother when the bandwidth is larger. When the bandwith is 0.01, the plot is more like broken-line, while when the bandwith is 1, the plot is like a normal distribution plot and centered at 0. The purple line indicates the normal distribution, and the black line is our estimation from kernal function.

```{r}
# Test the function with hand constructed datasets and bandwidths
set.seed(27)
n <- 500
x <- rnorm(n)
x_new <- rnorm(n)
h <- c(0.01, 0.05, 0.5, 1)

sapply(h, function(new_h) {
  plot(sort(x_new),  kern_density(x, new_h, sort(x_new)),type="l", main=paste("bandwidth =", new_h), 
       xlab= "test x", ylab="Kernel Density") 
  curve(dnorm(x), col='purple', lty = 4, add=TRUE)
})

```

# 2. CASL page 200, question 3.
Show that if $f$ and $g$ are both convex functions, then their sum must also be convex.

**answer**:
if $f(x)$ and $g(x)$ are convex functions then we will know that:
$$
f[\lambda x_1 + (1-\lambda)x_2] \le \lambda f(x_1) + (1-\lambda)f(x_2)
$$
and
$$
g[\lambda x_1 + (1-\lambda)x_2] \le \lambda g(x_1) + (1-\lambda)g(x_2)
$$
Suppose $h=f+g$, we have 
$$
\begin{aligned}
h[\lambda x_1 + (1-\lambda)x_2] &= f[\lambda x_1 + (1-\lambda)x_2]+g[\lambda x_1 + (1-\lambda)x_2] \\
&\le \lambda(f(x_1) + g(x_1)) + (1-\lambda)(f(x_2) + g(x_2)) \\
&= \lambda h(x_1) + (1-\lambda)h(x_2),
\end{aligned}
$$
Therefore $h$ as the sume of $f$ and $g$ is also a convex function.

# 3. CASL page 200, question 4.

Illustrate that the absolute value function is convex. Using the result from the previous exercise, show that the l1-norm is also convex.

**answer**:
L1-norm indicates the absolute value, therefore in this question, it asks me to prove that $f(x) = |x|$ is convex.

Let x, y $\in \mathbb{R}$, let $\alpha$, $\beta$ $\in \mathbb{R}_{>0}$ where $\alpha + \beta = 1$.
$$
\begin{aligned}
f(\alpha x + \beta y) = |\alpha x + \beta y| \leq |\alpha x| + |\beta y | = |\alpha| |x| + |\beta| |y| = \alpha |x| + \beta |y| = \alpha f(x) + \beta f(y)
\end{aligned}
$$

therefore,
$$
\begin{aligned}
f(\alpha x + \beta y) \leq \alpha f(x) + \beta f(y)
\end{aligned}
$$
Based on the above analysis, I proved that the l1-norm is also convex.

# 4. CASL page 200, question 5.
Prove that the elastic net objective function is convex using the results from the previous two exercises.

**answer**:
The elastic objective function is 

$$
f(b;\lambda, \alpha) = \frac{1}{2n} ||y-Xb||^2_2 + \lambda \left((1-\alpha) \frac{1}{2} ||b||_2^2 + \alpha||b||_1\right)
$$

Because $(y_i-X_{ij}b_j)^2$ is a convex function, according to question2, $||y-Xb||^2_2 = \sum_i \sum_j (y_i-X_{ij}b_j)^2$ is convex function; because $(b_j)^2$ is convex, then the sum $||b||_2^2 = \sum_j (b_j)^2$ is convex too.

According to question 3, the l-1 norm $||b||_1$ is a convex function as well. 

Back to the objective function, n>0, $\lambda > 0$ and $\alpha \in [0,1]$, so the sum of the three parts of convex function 
$$
\frac{1}{2n} ||y-Xb||^2_2 + \lambda \left((1-\alpha) \frac{1}{2} ||b||_2^2 + \alpha||b||_1\right)
$$
is a convex function. 

# 5. CASL page 200, question 6.
Find the KKT conditions for glmnet when 0 < $\alpha$ ≤ 1 and implement a lasso_reg_with_screening function that takes an $\alpha$ parameter.

**answer**:
```{r}
set.seed(27)
library(glmnet)
library(Matrix)
n <- 1000L
p <- 5000L
X <- matrix(rnorm(n * p), ncol = p)
beta <- c(seq(1, 0.1, length.out=(10L)), rep(0, p - 10L))
y <- X %*% beta + rnorm(n = n, sd = 0.15)
```

Then use glmnet to fit a lasso regression, check the KKT conditions to find the lambda value resulting in the minimum mean squared error). The result is only a very small portion of data resulting in invalid values.

```{r}
# glmnet
g_new<- cv.glmnet(X, y, alpha=1)
lambda.min <- g_new$lambda.min
beta_hat <- g_new$glmnet.g_new$beta[,which(g_new$lambda == g_new$lambda.min)]

# implement a lasso_reg_with_screening function
lasso_reg_with_screening <- function(X, y, b, lambda) {
  resids <- y - X %*% b
  res <- apply(X, 2, function(xj) crossprod(xj, resids)) / lambda / nrow(X)
  (b == 0) & (abs(res) >= 1)
}
```


