---
title: "Homework 4"
author: "Patty Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Homework 4 vignette}
-->

##1. CASL Number 2 in Exercises 5.8.

The Hessian matrix in Equation 5.19 is 
$$H(l) = X^t'diag(p(1-p)X)$$
where they are all inner product;

An ill-conditioned Hessian is where condition number is much larger than that of $X'X$. The condition number of a matrix is the ratio of its largest and smallest singular value.

Here I set a 2 by 2 matrix:
```{r}
X <- matrix(c(5, 0, 0, 1), 2, 2)
p <- c(0.5, 0.000001)
X
p
```

According to Equation 5.19, the logistic variation of Hessian matrix is 
$$X'diag(p(1-p))X$$
```{r}
#The condition number of this Hessian matrix is 
H = t(X)%*%diag(p*(1-p))%*%X
sv_1 <- svd(H)[["d"]]
condition_num_1<-max(sv_1)/min(sv_1)

#Then the condition number of $X'X$ is
sv_2 <- svd(t(X)%*%X)[["d"]]
condition_num_2<-max(sv_2)/min(sv_2)

#compare those two numbers
condition_num_1 - condition_num_2
```
The condition number of logistic variation Hessian matrix $X'diag(p(1-p))X$ is 6250006, while the condition number of $X'X$ is 25, 6250006 is way bigger than 25, therefore we would say that the logistic variabtion Hessian is ill-conditioned. But the linear Hessian is well conditioned.

##2. CASL Number 4 in Exercises 5.8.

Modify the function irwls_glm to include an l2-norm penalty on the regression vector for a fixed tuning parameter \lambda.

```{r}
# Args:
#     X: A numeric data matrix.
#     y: Response vector.
#     family: Instance of an R ‘family‘ object.
#     maxit: Integer maximum number of iterations.
#     tol: Numeric tolerance parameter.
#     lambda: a penalty term
#
# Returns:
#     Regression vector beta of length ncol(X).

irwls_glm_2 <- function(X, y, family, maxit=25, tol=1e-10, lambda){
  beta <- rep(0,ncol(X))
  for(j in 1:maxit){
    b_old <- beta
    eta   <- X %*% beta
    mu    <- family$linkinv(eta)
    mu_p  <- family$mu.eta(eta)
    z     <- eta + (y - mu) / mu_p
    W     <- as.numeric(mu_p^2 / family$variance(mu))
    #modify XtX term and add the 12-norm penalty
    XtX   <- crossprod(X, diag(W) %*% X) + diag(lambda, dim(XtX)[1])
    Xtz   <- crossprod(X, W * z)
    beta  <- solve(XtX, Xtz)
    if(sqrt(crossprod(beta - b_old)) < tol) break
  }
  beta
}
```





