---
title: "Ridge Regression Model MSE"
author: "Patty Zhang"
date: "10/14/2018"
output: html_document
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The linear model vignette}
-->

## The ridge regression function
The ridge regression function is to find an intelligent way of reducing the size of the prediction vector $\hat{\beta}$ while maintaining as much predictive power as possible. The ridge regression vector minimizes a combination of the sum of squares and the size of the regression vector, with the parameter $\lambda$ giving the relative balance between these two tasks. Here we are going to look at the effect on out-of-sample mean squared error as $\lambda$ varries and try to find the $\lambda$ which minimizes the mean squared error.

# Loading library and datasets
```{r}
library(bis557)
data(ridge_test)
data(ridge_train)

#one example
fit <- ridge_reg(Sepal.Length ~., lambda = 1, iris)
summary(fit)
```

# Looking at effect on the out-of-sample mean squared error as $\lambda$ varies
```{r}
# Predict values of ridge regression
# This is according to what we learnt in class
predict.ridge_reg <- function(object, ...){
  newdata <- list(...)[[1]]
  m <- model.matrix(object$form,newdata)
  m %*% object$coefficients
}

# Find the relationship between log(lambda) and mse
lambda_fit <- seq(0.1, 200, by=0.1)
mse = rep(NA, length(lambda_fit))
for (i in 1:length(lambda_fit)){
  fit <- ridge_reg(y~., lambda_fit[i], ridge_train)
  error <- ridge_test$y - predict(fit, ridge_test)
  mse[i] <- mean(error^2)
}

#plot mse with log(lambda) values, here we use log(lambda)
plot(mse~log(lambda_fit), xlab = "log(lambda)", ylab = "out-of-sample mean squared error values", type = "l")
#find the lambda value which minimizes out-of-sample mse
lambda_fit[which(mse == min(mse))]
```
